---
title: "RStudio_skills_assessments"
author: "Josue Tapia"
date: "11/6/2022"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#Limpiando el espacio de trabajo
rm(list=ls())
```

```{r}
#Libraries
library(tidyverse)
library(Hmisc)  #To use describe
library(ggplot2)
library(xtable)
library(coefplot)
```


## DATA

Upload data from my own github
```{r cars}
urlfile="https://raw.githubusercontent.com/JJTF/skills_assesstments/main/Data/wage2015_subsample_inference.csv" 
data<-read_csv(url(urlfile))
```
Describe data variables

```{r}
describe(data)
```

```{r}
str(data)
```
## variable graphs
After describe, we can know the type of data and choose a graph for each one.

Plot Wage
```{r}
hist(data$wage, xlab= "hourly wage", main="Wage distribution", breaks= 20)
```
Plot log wage
```{r}
hist(data$lwage, xlab= "hourly wage", main="Log Wage distribution", breaks= 20)
```

sex

Table
```{r}
table(data$sex)
```
Proportion
```{r}
round(prop.table(table(data$sex)), digits = 2)
```
Plot Sex
```{r}
pie(table(data$sex), labels = paste(c("Male","Female"), round(prop.table(table(data$sex))*100, digits = 1), "%", sep = "") , main = "Male & Female", col = rainbow(2))
legend("topleft", c("Male","Female"), cex = 0.6, fill = rainbow(2))
```



Show the frequency of variables shs,hsg,scl,clg,ad. We need it later to make a plot.
```{r}
sel_col <- c("shs","hsg","scl","clg","ad")
mod_frame <- apply(data[sel_col], 2, table)
print ("Frequencies")
colnames(mod_frame) <- c("Some High School","High School Graduate","Some College","College Graduate", "Advanced Degree")
print (mod_frame)

```



```{r}
variables <- c("shs","hsg","scl","clg","ad")
no <- c(5030,3894,3718,3514,4444)
yes <- c(120,1256,1432,1636,706)
df <- data.frame(variables, yes,no)
```


```{r}
require(tidyr)
df.long <- gather(df, variable,value, -variables)
```

Plot of variables: shs,hsg,scl,clg,ad
```{r}
ggplot(data = df.long, aes(x = variables, y = value, fill = variable)) +
  geom_col(position = position_dodge())
```


##Experience

```{r}
df_1<-data %>% 
  filter(sex == 1)%>%
  group_by(exp1)

df_2<- df_1 %>% group_by(exp1)%>% summarise(sex = n())
names(df_2)[2] <- "Female"

df_3<-data %>% 
  filter(sex == 0)%>%
  group_by(exp1)
df_4<- df_3 %>% group_by(exp1)%>% summarise(sex = n())
names(df_4)[2] <- 'Male'
data_1<- merge(df_2,df_4,by="exp1", all.y =TRUE)

```

Experience plot
Over here, i grouped by range the experience. The light blue represent female, and male the other color. In this sample we have more experience in  the range between 0 and 10.
```{r}
ggplot(data=data,aes(x=exp1 , fill=sex)) +
geom_bar(data=subset(data,sex==1)) +
geom_bar(data=subset(data,sex==0),aes(y=..count..*(-1))) +
labs(title="Experience by sex",x="Experience ranges",y="Male & Female")+
scale_y_continuous(breaks=seq(-100,100,400),labels=abs(seq(-100,100,400))) +
coord_flip()
```


```{r}
basic_model <- lwage ~(sex + exp1 + shs + hsg + mw + so + we + occ2 + ind2)
print(basic_model_regres<- lm(basic_model, data=data))
 
```

```{r}
flex_model <- lwage ~ sex + (exp1+ exp2+ exp3+ exp4+ shs + hsg + occ2+
ind2 + mw + so + we)^2
print(flex_model_regres <- lm(flex_model, data=data))
```
##Compare MSE and R2 adjusted
```{r}
basic_model_sum <- summary(basic_model_regres)
flex_model_sum <- summary(flex_model_regres)
n=5150
```

Rsquare basic model
```{r}
R2_bm <- basic_model_sum$r.squared
cat("Basic model - R2: ", R2_bm)
```
```{r}
R2_adj_bm <- basic_model_sum$adj.r.squared
cat("Basic model - R2 adjusted: ", R2_adj_bm)
```

Rsquare flexible model
```{r}
R2_fm <- flex_model_sum$r.squared
cat("Flexible model - R2: ", R2_fm)
```
```{r}
R2_adj_fm <- flex_model_sum$adj.r.squared
cat("Flexible model - R2 adjusted", R2_adj_fm)
```
MSE basic model
```{r}
MSE_bm <- mean(basic_model_sum$res^2)
cat("Basic model - MSE : ", MSE_bm)
```


```{r}
r_bm <- basic_model_sum$df[1] # number of regressors
MSE_adj_bm <- ( n/ (n-r_bm) )*MSE_bm
cat("Basic model - MSE adjusted ", MSE_adj_bm)
```


MSE flexible model
```{r}
MSE_fm <-mean(flex_model_sum$res^2)
cat("Flexible model - MSE :", MSE_fm)

```

```{r}
r_fm <- flex_model_sum$df[1]
MSE_fm_adj <- (n/(n-r_fm))*MSE_fm
cat("Flexible model - MSE adjusted:", MSE_fm_adj)
```


We can see at the table below that the flexible model has better performance than basic model. The last one has less $R^2$ and $R^2_{adjusted}$, even more $MSE$ and "$MSE_{adjusted}$". 
```{r}
table <- matrix(0, 2, 4)
table[1,1:4]   <- c(R2_bm,MSE_bm,R2_adj_bm,R2_adj_fm)
table[2,1:4]   <- c(R2_fm,MSE_fm,R2_adj_fm,MSE_fm_adj)

colnames(table)<- c("$R^2$","$MSE$","$R^2_{adjusted}$", "$MSE_{adjusted}$")
rownames(table)<- c("Basic model regres","Flexible model regres")
knitr::kable(table, "simple")
```



```{r}
#options(repr.plot.width=10,repr.plot.height=10)
#ggplot(data,aes(x=rownames(data), y="$R^2$"))+geom_point(size=1.5)+
#  geom_errorbar(aes(ymin=lower_bound))
```


